Neutron Star Mergers and their Impact on Second Generation Star Formation in the Early Universe

### 2023-07-14 NSM_mixing.tex
Updating Zr values, adding information about how to convert to [X/H].

To better represent the r-process metallicity, we define a new unit $Z_{\mathrm{r}} = 1.6 \times 10^{-10}$ which is scaled to the r-process abundances of the solar system from \citet{Goriely99}, normalized by the solar system H abundance. This means that a metallicity of 1 $Z_{\mathrm{r}}$ is equal to the r-process abundances within the solar system.

@@

\begin{table}
	\centering
	\begin{tabular}{lcccc} 
		\hline
		 & \# Pop III & M$_{\mathrm{Tot, P2}} [\Ms]$ & $f_{\mathrm{r}}$ & $f_{\mathrm{r, high}}$ \\
		\hline
		Fiducial & 2 & $10^{4.87}$ & 80\% & 80\% \\
		D & 2 & $10^{5.27}$ & 91\% & 63\% \\
		E & 1 & $10^{4.80}$ & 74\% & 74\% \\
		Original & 2 & $10^{4.52}$ & -- & -- \\
		\hline
	\end{tabular}
	\caption[Information about the stellar populations within the explosion energy variation runs.]{Information about the stellar populations within the explosion energy variation runs. For each run, we list the number of Pop III stars that form, the total Pop II stellar mass at the final redshift, the mass fraction of Pop II stars that are r-process enhanced ($> 10^{-8} Z_{\mathrm{r}}$), and the mass fraction of Pop II stars that are highly r-process enhanced ($> 10^{-1} Z_{\mathrm{r}}$).}
	\label{tab:eng_stellar}
\end{table}

@@

Figure \ref{fig:eng_NSM_deltat} shows the r-process metallicity of the Pop II stars that formed within the halo as a function of their creation time relative to the NSM event. We define stars to be ``highly r-process enhanced'' if they have a metallicity of $> 10^{-1} Z_{\mathrm{r}}$ and to be ``r-process enhanced'' if they have a metallicity of $> 10^{-8} Z_{\mathrm{r}}$. At $z = 10$, we find that runs D, Fiducial and E have Pop II mass fractions of 80\%, 91\% and 74\% being r-process enhanced, and Pop II mass fractions of 80\%, 63\%, and 74\% being highly r-process enhanced, respectively. The lowest explosion energy results in a lower percentage of highly r-process enhanced stars, but the highest percentage of r-process enhanced stars overall. As the energy increases, all Pop II stars that are enriched by the NSM are highly enhanced. This is due to the large amount of Pop II stars that form at late times near the center of the halo, where the majority of the r-process enhanced gas is contained. Figure \ref{fig:eng_NSM_deltat} also indicates that the Pop II stars that form closer to the NSM tend to have higher r-process metallicity fractions than if they were to form later on. We also see that the higher explosion energies produce Pop II stars with higher r-process metallicities, but less Pop II stars overall.

@@

\begin{table}
	\centering
	\begin{tabular}{lcccc} 
		\hline
		 & \# Pop III & M$_{\mathrm{Tot, P2}} [\Ms]$ & $f_{\mathrm{r}}$ & $f_{\mathrm{r, high}}$ \\
		\hline
		Fiducial & 2 & $10^{4.87}$ & 80\% & 80\% \\
		A & 6 & $10^{5.25}$ & 79\% & 50\% \\
		B & 3 & $10^{5.58}$ & 80\% & 24\% \\
		Original & 2 & $10^{4.52}$ & -- & -- \\
		\hline
	\end{tabular}
	\caption[Information about the stellar populations within the delay time variation runs.]{Information about the stellar populations within the delay time variation runs. For each run, we list the number of Pop III stars that form, the total Pop II stellar mass at the final redshift, the mass fraction of Pop II stars that are r-process enhanced ($> 10^{-8} Z_{\mathrm{r}}$), and the mass fraction of Pop II stars that are highly r-process enhanced ($> 10^{-1} Z_{\mathrm{r}}$).}
	\label{tab:time_stellar}
\end{table}

@@

Figure \ref{fig:time_NSM_deltat} shows the r-process metallicity for Pop II stars as a function of their creation time relative to when the NSM occurred. In this plot, there are obvious groups of stars that come from the different runs. Run A, with the shortest delay time, forms Pop II stars at later times compared to when the NSM occurs, and at both high and low r-process metallicities. Run B, with the 30 Myr delay time, forms the highest amount of Pop II stars, but with relatively lower r-process metallicities overall. The Fiducial run forms Pop II stars the soonest after the NSM, which occurs simply because the delay time is longer, and with higher r-process metallicities. In this run, there hasn't been enough time for the material to escape the halo before Pop II Stars form, therefore, higher r-process metallicities are expected. We find that run A results in a 79\% and 50\% of the Pop II mass fraction being r-process enhanced and highly r-process enhanced, run B resulting in 80\% and 24\%, and run Fiducial resulting in 80\% and 80\% of the Pop II stars being r-process enhanced and highly r-process enhanced, respectively. This indicates that the NSM is in general enhancing a similar percentage of Pop II stars with r-process material, but the shorter delay times result in a smaller percentage of Pop II stars being highly r-process enhanced.

@@

We have run a suite of simulations studying the impact of Pop III remnants, in the form of NSMs, on the second generation of star formation in the universe. We ran one simulation without a NSM, and five simulations varying the explosion energy and the delay time of the NSM in order to see how these variations affect the r-process abundances of the next generation of stars. In our case study, we find that:
\begin{itemize}
	\item NSMs originating from Pop III stars do lead to a substantial fraction of highly r-process enhanced Pop II stars.
	\item When the explosion energy is high, 74\% of stars are highly r-process enhanced, versus when the explosion energy is lower, only 63\% are highly r-process enhanced. 
	\item The lower explosion energy leads to a higher fraction of Pop II stars being r-process enhanced overall, but at lower metallicities in general. The higher explosion energy leads to a smaller fraction of r-process enhanced stars, but all that are enhanced are highly enhanced. 
	\item The longer time delays lead to higher mass fractions of highly r-process enhanced stars, due to the NSM occurring closer to the period of Pop II star formation. 
	\item The shorter time delays still lead to a high mass fraction of r-process enhanced stars, but a smaller fraction of highly r-process enhanced stars.
	\item The half-mass radius significantly decreases in each NSM run due to the large number of Pop II stars that form and cluster near the center of the halo at the end of the simulation. 
	\item Star formation and feedback that occur in and around the halo heavily govern the dynamics and details of star formation in these runs. A larger sample size of galaxies with NSMs is needed in order to more fully understand the impact of a NSM on the second generation of stars. 
\end{itemize}

>>>>

To better represent the r-process metallicity, we define a new unit $Z_{\mathrm{r}} = 0.01384$ which is scaled to the r-process abundances of the solar system from \citet{Goriely99}, normalized by the solar system Si abundance. This means that a metallicity of 1 $Z_{\mathrm{r}}$ is equal to the r-process abundances within the solar system. To convert from $Z_{\mathrm{r}}$ to an abundance ratio of [X/H] for some element X, one can use $[X/H] = [X/H]_{\mathrm{r}} + A(X)_{\odot, r} - A(X)_{\odot}$ to normalize to an r-process solar composition \cite[see Step \#3 in Table 2 of][for an excellent description of metallicity conversions]{Hinkel22}. Here, $[X/H]_{\mathrm{r}} = v_{x} + \mathrm{log}(Z_{\mathrm{r}})$, where $v_{x}$ is open to the reader, depending on their choice of NSM ejecta yields. This makes $[X/H] = v_{x} + \mathrm{log}(Z_{\mathrm{r}}) + A(X)_{\odot, r} - A(X)_{\odot}$. To compare with results from \citet{Frebel23}, we put this in terms of abundances from \citet{Lodders09} (L09) and \citet{Roederer22} (R22) as collected by \citet{Frebel23} (see their Table 2). In terms of these values, $[X/H] = v_{x} + \mathrm{log}(Z_{\mathrm{r}}) + \mathrm{log} \epsilon_{\mathrm{R22}} (X) - \mathrm{log} \epsilon_{\mathrm{L09}} (X)$. For Ba, the difference between the abundances from \citet{Frebel23} is -0.92 and for Eu, the difference between the abundances is -0.12.

@@

\begin{table}
	\centering
	\begin{tabular}{lcccc} 
		\hline
		 & \# Pop III & M$_{\mathrm{Tot, P2}} [\Ms]$ & $f_{\mathrm{r}}$ & $f_{\mathrm{r, high}}$ \\
		\hline
		Fiducial & 2 & $10^{4.87}$ & 80\% & 64\% \\
		D & 2 & $10^{5.27}$ & 80\% & 14\% \\
		E & 1 & $10^{4.80}$ & 74\% & 72\% \\
		Original & 2 & $10^{4.52}$ & -- & -- \\
		\hline
	\end{tabular}
	\caption[Information about the stellar populations within the explosion energy variation runs.]{Information about the stellar populations within the explosion energy variation runs. For each run, we list the number of Pop III stars that form, the total Pop II stellar mass at the final redshift, the mass fraction of Pop II stars that are r-process enhanced ($> 10^{-6} Z_{\mathrm{r}}$), and the mass fraction of Pop II stars that are highly r-process enhanced ($> 10^{-1} Z_{\mathrm{r}}$).}
	\label{tab:eng_stellar}
\end{table}

@@

Figure \ref{fig:eng_NSM_deltat} shows the r-process metallicity of the Pop II stars that formed within the halo as a function of their creation time relative to the NSM event. We define stars to be ``highly r-process enhanced'' if they have a metallicity of $> 10^{-1} Z_{\mathrm{r}}$ and to be ``r-process enhanced'' if they have a metallicity of $> 10^{-6} Z_{\mathrm{r}}$. At $z = 10$, we find that runs D, Fiducial and E have Pop II mass fractions of 80\%, 80\% and 74\% being r-process enhanced, and Pop II mass fractions of 14\%, 63\%, and 72\% being highly r-process enhanced, respectively. The lowest explosion energy results in a lower percentage of highly r-process enhanced stars, but a similar percentage of r-process enhanced stars compared to the higher explosion energy runs. As the energy increases, all Pop II stars that are enriched by the NSM are highly enhanced. This is due to the large amount of Pop II stars that form at late times near the center of the halo, where the majority of the r-process enhanced gas is contained. Figure \ref{fig:eng_NSM_deltat} also indicates that the Pop II stars that form closer to the NSM tend to have higher r-process metallicity fractions than if they were to form later on. We also see that the higher explosion energies produce Pop II stars with higher r-process metallicities, but less Pop II stars overall.

@@

\begin{table}
	\centering
	\begin{tabular}{lcccc} 
		\hline
		 & \# Pop III & M$_{\mathrm{Tot, P2}} [\Ms]$ & $f_{\mathrm{r}}$ & $f_{\mathrm{r, high}}$ \\
		\hline
		Fiducial & 2 & $10^{4.87}$ & 80\% & 64\% \\
		A & 6 & $10^{5.25}$ & 70\% & 5\% \\
		B & 3 & $10^{5.58}$ & 74\% & 0.2\% \\
		Original & 2 & $10^{4.52}$ & -- & -- \\
		\hline
	\end{tabular}
	\caption[Information about the stellar populations within the delay time variation runs.]{Information about the stellar populations within the delay time variation runs. For each run, we list the number of Pop III stars that form, the total Pop II stellar mass at the final redshift, the mass fraction of Pop II stars that are r-process enhanced ($> 10^{-6} Z_{\mathrm{r}}$), and the mass fraction of Pop II stars that are highly r-process enhanced ($> 10^{-1} Z_{\mathrm{r}}$).}
	\label{tab:time_stellar}
\end{table}

@@

Figure \ref{fig:time_NSM_deltat} shows the r-process metallicity for Pop II stars as a function of their creation time relative to when the NSM occurred. In this plot, there are obvious groups of stars that come from the different runs. Run A, with the shortest delay time, forms Pop II stars at later times compared to when the NSM occurs, and at both high and low r-process metallicities. Run B, with the 30 Myr delay time, forms the highest amount of Pop II stars, but with relatively lower r-process metallicities overall. The Fiducial run forms Pop II stars the soonest after the NSM, which occurs simply because the delay time is longer, and with higher r-process metallicities. In this run, there hasn't been enough time for the material to escape the halo before Pop II Stars form, therefore, higher r-process metallicities are expected. We find that run A results in a 70\% and 5\% of the Pop II mass fraction being r-process enhanced and highly r-process enhanced, run B resulting in 74\% and 0.02\%, and run Fiducial resulting in 80\% and 64\% of the Pop II stars being r-process enhanced and highly r-process enhanced, respectively. This indicates that the NSM is in general enhancing a similar percentage of Pop II stars with r-process material, but the shorter delay times result in a much smaller percentage of Pop II stars being highly r-process enhanced.

@@

We have run a suite of simulations studying the impact of Pop III remnants, in the form of NSMs, on the second generation of star formation in the universe. We ran one simulation without a NSM, and five simulations varying the explosion energy and the delay time of the NSM in order to see how these variations affect the r-process abundances of the next generation of stars. In our case study, we find that:
\begin{itemize}
	\item NSMs originating from Pop III stars do lead to a substantial fraction of highly r-process enhanced Pop II stars.
	\item When the explosion energy is high, 73\% of stars are highly r-process enhanced, versus when the explosion energy is lower, only 14\% are highly r-process enhanced. 
	\item The lower explosion energy leads to a higher fraction of Pop II stars being r-process enhanced overall, but at lower metallicities in general. The higher explosion energy leads to a smaller fraction of r-process enhanced stars, but almost all that are enhanced are highly enhanced. 
	\item The longer time delays lead to higher mass fractions of highly r-process enhanced stars, due to the NSM occurring closer to the period of Pop II star formation. 
	\item The shorter time delays still lead to a high mass fraction of r-process enhanced stars, but a much smaller fraction of highly r-process enhanced stars.
	\item The half-mass radius significantly decreases in each NSM run due to the large number of Pop II stars that form and cluster near the center of the halo at the end of the simulation. 
	\item Star formation and feedback that occur in and around the halo heavily govern the dynamics and details of star formation in these runs. A larger sample size of galaxies with NSMs is needed in order to more fully understand the impact of a NSM on the second generation of stars. 
\end{itemize}
